SFJ: An implementation of Semantic Featherweight Java

### 2020-04-27 conference/camera ready/background.tex
giving space

\subsection{Types}
\label{sec:types}
The syntax of types $\tau$ is defined by the following grammar:
$$
\begin{array}{llll}
\tau & ::= & \alpha \ |\ \mu
& \mbox{{\em Types}}
\\
\alpha & ::=  & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha\ \synwedge\ \alpha \ |\ \synneg\ \alpha
\qquad
& \mbox{{\em Field types} ($\alpha $-types)}
\\
\mu & ::=  & \alpha \to \alpha \ |\ \mu\ \synwedge\ \mu \ |\ \synneg\ \mu
& \mbox{{\em Method types} ($\mu $-types)}
\end{array}
$$
The $\alpha$-types are used to type fields and the $\mu$-types are used to type methods.
Type $\zero$ is the empty type.
Type $\btypes$ denotes the \emph{basic} types, such as integers, booleans, etc.
{Record} types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels, are used to type objects.
Arrow types $\alpha \to \alpha$ are used to type methods.

@@

\subsection{Terms}
\label{sec:terms}
The syntax of terms is defined by the following grammar and is based on the standard syntax of terms in FJ \cite{featherweight}:
\begin{align*}
     & \mbox{\textit{Class declaration}}  & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }                         \\
     & \mbox{\textit{Constructor}}        & K \;::=\;   & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \} \\
     & \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \}                             \\
     & \mbox{\textit{Expressions}}        & e \; ::=\;  & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
\end{align*}
We assume an infinite set of names, with some special names:
$\mathit{Object}$ denotes the root class, $\synthis$ denotes the current object and $\synsuper$ denotes the parent object.
We let  $A, B,\ldots$ range over classes; $a, b,\ldots$ over fields; $m, n,\ldots$ over methods and $x, y, z, \ldots$ over variables.

>>>>

\subsection{Types}
\label{sec:types}
The syntax of types $\tau$ is defined by the following grammar:
$$
\begin{array}{llll}
\tau & ::= & \alpha \ |\ \mu
& \mbox{{\em Types}}
\\
\alpha & ::=  & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha\ \synwedge\ \alpha \ |\ \synneg\ \alpha
\qquad
& \mbox{{\em Field types} ($\alpha $-types)}
\\
\mu & ::=  & \alpha \to \alpha \ |\ \mu\ \synwedge\ \mu \ |\ \synneg\ \mu
& \mbox{{\em Method types} ($\mu $-types)}
\end{array}
$$

The $\alpha$-types are used to type fields and the $\mu$-types are used to type methods.
Type $\zero$ is the empty type.
Type $\btypes$ denotes the \emph{basic} types, such as integers, booleans, etc.
{Record} types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels, are used to type objects.
Arrow types $\alpha \to \alpha$ are used to type methods.

@@

\subsection{Terms}
\label{sec:terms}
The syntax of terms is defined by the following grammar and is based on the standard syntax of terms in FJ \cite{featherweight}:
\begin{align*}
     & \mbox{\textit{Class declaration}}  & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }                         \\
     & \mbox{\textit{Constructor}}        & K \;::=\;   & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \} \\
     & \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \}                             \\
     & \mbox{\textit{Expressions}}        & e \; ::=\;  & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
\end{align*}

We assume an infinite set of names, with some special names:
$\mathit{Object}$ denotes the root class, $\synthis$ denotes the current object and $\synsuper$ denotes the parent object.
We let  $A, B,\ldots$ range over classes; $a, b,\ldots$ over fields; $m, n,\ldots$ over methods and $x, y, z, \ldots$ over variables.

### 2020-04-27 conference/camera ready/paper.tex
giving space

    Consider the set of polygons, such as triangles, squares and rhombuses given by a class hierarchy.
    We want to define a method \emph{diagonal} that takes a polygon and returns the length of its longest diagonal.
    This method makes sense only if the polygon passed to it has at least four sides, hence triangles are excluded.
    In Java this could be implemented in the following ways:
    \begin{equation*}
        \begin{array}{l}
            \synclass Polygon \ \{ \ldots\}
            \\
            \syndecl{Triangle}{Polygon} {\ldots}
            \\
            \syndecl{Other\_Polygons}{Polygon} {
                \\
                \qquad \mathbf{double}\ diagonal(Other\_Polygons \ shape)\ \{\ldots \}
                \\
                \hspace{3.5cm} \ldots
                \\
            }
            \\
            \syndecl{Square}{Other\_Polygons}{\ldots}
            \\
            \syndecl{Rhombus}{Other\_Polygons} {\ldots}
        \end{array}
    \end{equation*}
    Or by means of an interface \emph{Diagonal}:
    \begin{equation*}
        \begin{array}{l}
            public \ \syninterface Diagonal\ \{
            \\
            \qquad \mathbf{double}\ diagonal(Polygon \ shape);
            \\
            \}
            \\
            \synclass Polygon \ \{\ldots\}
            \\
            \syndecl{Triangle}{Polygon} {\ldots}
            \\
            \synidecl{Square}{Polygon} \synimpl Diagonal \ \{\ldots\}
            \\
            \synidecl{Rhombus}{Polygon} \synimpl Diagonal \ \{\ldots\}
            \\
            \hspace{2.5cm} // \text{ other polygons } \ldots
        \end{array}
    \end{equation*}
    Now, suppose our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figures extend \emph{Polygon}, which is how one would naturally define the set of polygons.
    Suppose the class hierarchy is given and is part of legacy code, which cannot be changed.
    Then again, a natural way to implement this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for example, inside a \textbf{try-catch} construct.
    Then, an exception would be thrown at run time, if the argument passed to the method is a triangle.

    We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static type-checking is required and we implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}, thus allowing the type-checker to check at compile time the restrictions on types:
    \begin{equation*}
        \begin{array}{l}
            \synclass Polygon \ \{\ldots\}
            \\
            \syndecl{Triangle}{Polygon} {\ldots}
            \\
            \syndecl{Square}{Polygon} {\ldots}
            \\
            \syndecl{Rhombus}{Polygon} {\ldots}
            \\
            \hspace{2.5cm} \ldots
            \\
            \synclass Diagonal \ \{
            \\
            %\qquad Diagonal ( )\   \{\ super();\ \}
            \hspace{2.5cm} \ldots \\
            \qquad \mathbf{double}\ diagonal((\textit{Polygon} \textbf{ and not }  \textit{Triangle}) \ shape)\{\ldots\}
            \\
            \}
        \end{array}
    \end{equation*}
    We can now call \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is \textbf{not} a \emph{Triangle}, then the method computes and returns the length of its longest diagonal; otherwise, there will be a type error at compile time.
\end{example}

@@

Running the parser on an SFJ program returns an abstract syntax tree (AST) of that program.
When visiting the AST, we check if the program is well-formed, following the intuition at the end of \autoref{sec:background}.
We mark any classes containing only fields typed with basic types as \emph{resolved}, otherwise we mark them as \emph{unresolved}.
Using this information, \autoref{alg:types} checks if the type definitions in a program are valid, namely if they are finite trees whose leaves are basic types with no cycles.
%
The algorithm returns $True$ only if all the types in the program are resolved, otherwise it returns $False$, meaning there is at least one type definition which is invalid and contains a cycle.

@@

\subsection{Building Semantic Subtyping for SFJ}
\label{sec:subtypingAlgo}
If \autoref{alg:types} returns $True$, meaning all type definitions in a program are valid, we can then build the semantic subtyping.
Leveraging the interpretation of types as sets of values to define semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in SFJ we keep track of the semantic subtyping relation by defining a map from a type to the set of its subtypes, satisfying the property that the set of values of a subtype is included in the set of values of the type.
We start with basic types and let $Universe$ be a supertype of all types.
The full mapping for basic types is defined in \autoref{eq:initialSubtype}.
\begin{equation}
    \label{eq:initialSubtype}
    \begin{array}{llllll}
        & Double   & = \left\{Double, Float, Int, Short, Byte\right\} &  & Float & = \left\{Float, Short, Byte\right\} \\
        & Long     & =    \left\{Long, Int, Short, Byte\right\}       &  & Int   & = \left\{Int, Short, Byte\right\}   \\
        & Short    & =    \left\{Short, Byte\right\}                  &  & Byte  & = \left\{Byte\right\}               \\
        & Boolean  & =    \left\{Boolean\right\}                      &  & Void  & = \left\{Void\right\}               \\
        & Universe & = \{Double, Float, Long, Int,                                                                     \\
        &          & \quad Short, Byte, Boolean, Void\}
   \end{array}
\end{equation}
Note that \emph{Int} is not a subtype of \emph{Float} as a 32-bit \emph{float} cannot represent the whole set of 32-bit \emph{integer} values accurately and therefore \emph{Int} is not fully set-contained in \emph{Float}, however this is not the case for \emph{Int} and \emph{Double}.
Similarly, \emph{Long} is not a subtype of \emph{Double}.

\autoref{alg:subtyping} builds the semantic subtyping relation for all class types of an SFJ program by calling the function \textit{generateRelation}.
Given that classes are valid type definitions by \autoref{alg:types}, we are guaranteed that \autoref{alg:subtyping} will terminate.
The semantic subtyping generated by \autoref{alg:subtyping} is a preorder: it is reflexive and transitive. This is also illustrated by \autoref{eq:initialSubtype}.

@@

Let us illustrate the semantic subtyping algorithm on our \emph{Polygons} given in \autoref{polygons}.
\autoref{alg:subtyping} generates the subtyping relation given in \autoref{eq:sugtypePolygons}, together with the subtyping relation for basic types, omitted here and defined in \autoref{eq:initialSubtype}. Notice that the mapping for $Universe$ is extended with the new types for polygons.
\begin{equation}
    \label{eq:sugtypePolygons}
    \begin{array}{llllll}
         & Polygon  & = \left\{Polygon, Triangle, Square, Rhombus \right\} &  & Triangle & = \left\{Triangle\right\}           \\
         & Square   & = \left\{Square\right\}                              &  & Rhombus  & = \left\{Rhombus\right\}            \\
         & Universe & = \{Double, Float, Long, Int, Short, Byte                                                              &  & Diagonal & = \left\{Diagonal\right\}          \\
         &          & \quad Boolean, Void, Polygon, Square                                                                     \\
         &          & \quad Square, Rhombus, Diagonal\}
    \end{array}
\end{equation}
Recall the method \emph{diagonal} in class \emph{Diagonal}, with signature
$$
\mathbf{double}\ diagonal((\textit{Polygon} \textbf{ and not }  \textit{Triangle}) \ shape)
$$
The result of the set operation on its parameter type gives the following set of polygons:
$$
Polygon \textbf{ and not } Triangle = \left\{Polygon, Square, Rhombus \right\}
$$
In order to define the \textbf{not} $Triangle$ type we need the $Universe$ type so that we can define it as $Universe\ \mathbf{\setminus}\ Triangle$. Then, the \textbf{and} connective is the intersection of sets of $Polygon$ with \textbf{not} $Triangle$.

@@

On the other hand, while nominal subtyping can avoid the above problem, it can introduce others and in particular, a developer can define an overridden method to perform the opposite logic to what the super class is expecting, as illustrated by the following classes in Java:
\begin{equation*}
    \begin{array}{ll}
        \syndecl{A}{\mathit{Object}} {
         &
            \qquad
            \syndecl{B}{A} {
                \\
               \qquad \qquad\ldots \qquad
         &
                \qquad\qquad\qquad \ldots
                \\
                \quad \textbf{int}\ n;\qquad
         &
                \qquad\quad \textbf{int}\ length()\{\ \synret{-n};\ \}
                \\
                \quad \textbf{int}\ length()\{\ \synret{n};\ \}\qquad
         &
                \qquad}
                \\   }
    \end{array}
\end{equation*}
Both approaches have their pros and cons, and they leave an expectation on the developer to use the logic behind subtyping correctly when writing code.
The integration of both subtyping approaches in SFJ gives the developer the freedom to choose the most suitable subtyping relation to use for a given task.

@@

\subsection{Methods in SFJ}
\label{sec:methods}
\subsubsection{On multimethods}
Since FJ is a core language, some features of the full Java are removed, such as overloading methods.
In our framework, by leveraging the expressivity of boolean connectives and semantic subtyping, we are able to restore overloading, among other features \cite[\S 8.4]{Dardha2017}.
We can thus model \emph{multimethods}, \cite{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
As an example, taken from Dardha \emph{et al.}  \cite{Dardha2013,Dardha2017}, consider the following class declarations:
$$
\begin{array}{ll}
\syndecl{A}{\mathit{Object}} {
&
\qquad
\syndecl{B}{A} {
\\
	%\qquad\quad \ldots
%&
	%\qquad\qquad\quad \ldots
%\\
	\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
&
	\qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
\\
}
&
\qquad }
\end{array}
$$
Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in class $A$.
However, because class $B$ extends class $A$,  {\it length} has type $(\mathbf{string} \rightarrow \mathbf{int})\ \synwedge\ (\mathbf{int} \rightarrow \mathbf{int})$ in class $B$,
which can be simplified to $(\mathbf{string}\ \synvee\ \mathbf{int}) \rightarrow \mathbf{int}$.

\subsubsection{Method types}
Let us illustrate the method types given in \autoref{sec:types} via an alternative implementation of the class $Diagonal$ at the end of \autoref{polygons}.
$$
\begin{array}{l}
    \synclass Diagonal \ \{
    \\
    \qquad\qquad\qquad \ldots \\
    \qquad \textbf{double}\ diagonal((diagonal: \textbf{void} \rightarrow \textbf{double})\ shape)\\
    \hfill
    \{\ \synret\ shape.diagonal();\ \}
    \\
    \}
\end{array}
$$
We define the type of the (outside) $diagonal$ method as accepting any type and its subtypes implementing the (inside) \emph{diagonal} method with type signature \textbf{void} to \textbf{double}.

>>>>

    Consider the set of polygons, such as triangles, squares and rhombuses given by a class hierarchy.
    We want to define a method \emph{diagonal} that takes a polygon and returns the length of its longest diagonal.
    This method makes sense only if the polygon passed to it has at least four sides, hence triangles are excluded.
    In Java this could be implemented in the following ways:
    \begin{equation*}
        \begin{array}{l}
            \synclass Polygon \ \{ \ldots\}
            \\
            \syndecl{Triangle}{Polygon} {\ldots}
            \\
            \syndecl{Other\_Polygons}{Polygon} {
                \\
                \qquad \mathbf{double}\ diagonal(Other\_Polygons \ shape)\ \{\ldots \}
                \\
                \hspace{3.5cm} \ldots
                \\
            }
            \\
            \syndecl{Square}{Other\_Polygons}{\ldots}
            \\
            \syndecl{Rhombus}{Other\_Polygons} {\ldots}
        \end{array}
    \end{equation*}
    Or by means of an interface \emph{Diagonal}:
    \begin{equation*}
        \begin{array}{l}
            public \ \syninterface Diagonal\ \{
            \\
            \qquad \mathbf{double}\ diagonal(Polygon \ shape);
            \\
            \}
            \\
            \synclass Polygon \ \{\ldots\}
            \\
            \syndecl{Triangle}{Polygon} {\ldots}
            \\
            \synidecl{Square}{Polygon} \synimpl Diagonal \ \{\ldots\}
            \\
            \synidecl{Rhombus}{Polygon} \synimpl Diagonal \ \{\ldots\}
            \\
            \hspace{2.5cm} // \text{ other polygons } \ldots
        \end{array}
    \end{equation*}

    Now, suppose our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figures extend \emph{Polygon}, which is how one would naturally define the set of polygons.
    Suppose the class hierarchy is given and is part of legacy code, which cannot be changed.
    Then again, a natural way to implement this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for example, inside a \textbf{try-catch} construct.
    Then, an exception would be thrown at run time, if the argument passed to the method is a triangle.

    We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static type-checking is required and we implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}, thus allowing the type-checker to check at compile time the restrictions on types:
    \begin{equation*}
        \begin{array}{l}
            \synclass Polygon \ \{\ldots\}
            \\
            \syndecl{Triangle}{Polygon} {\ldots}
            \\
            \syndecl{Square}{Polygon} {\ldots}
            \\
            \syndecl{Rhombus}{Polygon} {\ldots}
            \\
            \hspace{2.5cm} \ldots
            \\
            \synclass Diagonal \ \{
            \\
            %\qquad Diagonal ( )\   \{\ super();\ \}
            \hspace{2.5cm} \ldots \\
            \qquad \mathbf{double}\ diagonal((\textit{Polygon} \textbf{ and not }  \textit{Triangle}) \ shape)\{\ldots\}
            \\
            \}
        \end{array}
    \end{equation*}

    We can now call \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is \textbf{not} a \emph{Triangle}, then the method computes and returns the length of its longest diagonal; otherwise, there will be a type error at compile time.
\end{example}

@@

Running the parser on an SFJ program returns an abstract syntax tree (AST) of that program.
When visiting the AST, we check if the program is well-formed, following the intuition at the end of \autoref{sec:background}.
We mark any classes containing only fields typed with basic types as \emph{resolved}, otherwise we mark them as \emph{unresolved}.
Using this information, \autoref{alg:types} checks if the type definitions in a program are valid, namely if they are finite trees whose leaves are basic types with no cycles.
%
The algorithm returns $True$ only if all the types in the SFJ program are resolved, otherwise it returns $False$, meaning there is at least one type definition which is invalid and contains a cycle.

@@

\subsection{Building Semantic Subtyping for SFJ}
\label{sec:subtypingAlgo}
If \autoref{alg:types} returns $True$, meaning all type definitions in a program are valid, we can then build the semantic subtyping.
Leveraging the interpretation of types as sets of values to define semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in SFJ we keep track of the semantic subtyping relation by defining a map from a type to the set of its subtypes, satisfying the property that the set of values of a subtype is included in the set of values of the type.

We start with basic types and let $Universe$ be a supertype of all types.
The full mapping for basic types is defined in \autoref{eq:initialSubtype}.
\begin{equation}
    \label{eq:initialSubtype}
    \begin{array}{llllll}
        & Double   & = \left\{Double, Float, Int, Short, Byte\right\} &  & Float & = \left\{Float, Short, Byte\right\} \\
        & Long     & =    \left\{Long, Int, Short, Byte\right\}       &  & Int   & = \left\{Int, Short, Byte\right\}   \\
        & Short    & =    \left\{Short, Byte\right\}                  &  & Byte  & = \left\{Byte\right\}               \\
        & Boolean  & =    \left\{Boolean\right\}                      &  & Void  & = \left\{Void\right\}               \\
        & Universe & = \{Double, Float, Long, Int,                                                                     \\
        &          & \quad Short, Byte, Boolean, Void\}
   \end{array}
\end{equation}
Note that \emph{Int} is not a subtype of \emph{Float} as a 32-bit \emph{float} cannot represent the whole set of 32-bit \emph{integer} values accurately and therefore \emph{Int} is not fully set-contained in \emph{Float}, however this is not the case for \emph{Int} and \emph{Double}.
Similarly, \emph{Long} is not a subtype of \emph{Double}.

\autoref{alg:subtyping} builds the semantic subtyping relation for all class types of an SFJ program by calling the function \textit{generateRelation}.
Given that classes are valid type definitions by \autoref{alg:types}, we are guaranteed that \autoref{alg:subtyping} will terminate.

The semantic subtyping generated by \autoref{alg:subtyping} is a preorder: it is reflexive and transitive. This is also illustrated by \autoref{eq:initialSubtype}.

@@

Let us illustrate the semantic subtyping algorithm on our \emph{Polygons} given in \autoref{polygons}.
\autoref{alg:subtyping} generates the subtyping relation given in \autoref{eq:sugtypePolygons}, together with the subtyping relation for basic types, omitted here and defined in \autoref{eq:initialSubtype}. Notice that the mapping for $Universe$ is extended with the new types for polygons.
\begin{equation}
    \label{eq:sugtypePolygons}
    \begin{array}{llllll}
         & Polygon  & = \left\{Polygon, Triangle, Square, Rhombus \right\} &  & Triangle & = \left\{Triangle\right\}           \\
         & Square   & = \left\{Square\right\}                              &  & Rhombus  & = \left\{Rhombus\right\}            \\
         & Universe & = \{Double, Float, Long, Int, Short, Byte                                                              &  & Diagonal & = \left\{Diagonal\right\}          \\
         &          & \quad Boolean, Void, Polygon, Square                                                                     \\
         &          & \quad Square, Rhombus, Diagonal\}
    \end{array}
\end{equation}

Recall the method \emph{diagonal} in class \emph{Diagonal}, with signature
$$
\mathbf{double}\ diagonal((\textit{Polygon} \textbf{ and not }  \textit{Triangle}) \ shape)
$$
The result of the set operation on its parameter type gives the following set of polygons:
$$
Polygon \textbf{ and not } Triangle = \left\{Polygon, Square, Rhombus \right\}
$$
In order to define the \textbf{not} $Triangle$ type we need the $Universe$ type so that we can define it as $Universe\ \mathbf{\setminus}\ Triangle$. Then, the \textbf{and} connective is the intersection of sets of $Polygon$ with \textbf{not} $Triangle$.

@@

On the other hand, while nominal subtyping can avoid the above problem, it can introduce others and in particular, a developer can define an overridden method to perform the opposite logic to what the super class is expecting, as illustrated by the following classes in Java:
\begin{equation*}
    \begin{array}{ll}
        \syndecl{A}{\mathit{Object}} {
         &
            \qquad
            \syndecl{B}{A} {
                \\
               \qquad \qquad\ldots \qquad
         &
                \qquad\qquad\qquad \ldots
                \\
                \quad \textbf{int}\ n;\qquad
         &
                \qquad\quad \textbf{int}\ length()\{\ \synret{-n};\ \}
                \\
                \quad \textbf{int}\ length()\{\ \synret{n};\ \}\qquad
         &
                \qquad}
                \\   }
    \end{array}
\end{equation*}

Both approaches have their pros and cons, and they leave an expectation on the developer to use the logic behind subtyping correctly when writing code.
The integration of both subtyping approaches in SFJ gives the developer the freedom to choose the most suitable subtyping relation to use for a given task.

@@

\subsection{Methods in SFJ}
\label{sec:methods}
\subsubsection{On multimethods}
Since FJ is a core language, some features of the full Java are removed, such as overloading methods.
In our framework, by leveraging the expressivity of boolean connectives and semantic subtyping, we are able to restore overloading, among other features \cite[\S 8.4]{Dardha2017}.
We can thus model \emph{multimethods}, \cite{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
As an example, taken from Dardha \emph{et al.}  \cite{Dardha2013,Dardha2017}, consider the following class declarations:
$$
\begin{array}{ll}
\syndecl{A}{\mathit{Object}} {
&
\qquad
\syndecl{B}{A} {
\\
	%\qquad\quad \ldots
%&
	%\qquad\qquad\quad \ldots
%\\
	\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
&
	\qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
\\
}
&
\qquad }
\end{array}
$$

Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in class $A$.
However, because class $B$ extends class $A$,  {\it length} has type $(\mathbf{string} \rightarrow \mathbf{int})\ \synwedge\ (\mathbf{int} \rightarrow \mathbf{int})$ in class $B$,
which can be simplified to $(\mathbf{string}\ \synvee\ \mathbf{int}) \rightarrow \mathbf{int}$.

\subsubsection{Method types}
Let us illustrate the method types given in \autoref{sec:types} via an alternative implementation of the class $Diagonal$ at the end of \autoref{polygons}.
$$
\begin{array}{l}
    \synclass Diagonal \ \{
    \\
    \qquad\qquad\qquad \ldots \\
    \qquad \textbf{double}\ diagonal((diagonal: \textbf{void} \rightarrow \textbf{double})\ shape)\\
    \hfill
    \{\ \synret\ shape.diagonal();\ \}
    \\
    \}
\end{array}
$$

We define the type of the (outside) $diagonal$ method as accepting any type and its subtypes implementing the (inside) \emph{diagonal} method with type signature \textbf{void} to \textbf{double}.
